---
title: "ProjectOne"
output: html_document
---


# Data Project One

## 100 points
## DUE Sept 15 at 11:59 PM

Below are the questions for the first data practical assignment. This project uses the "FossilAnts.csv" file, located in the data directory for the project. The point value of each question is denoted next to it. A blank cell is below each for your answer; feel free to create more blank cells as needed.

0. (5 pts) Create a directory called `projects`, and in it, a subdirectory called `project_one`. Use `download.files` to get [this file](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/project_one.Rmd), and [this one](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/projects/project_one/FossilAnts.csv) and save them to the `project_one` directory.


<!-- 1. 5 pts. Import the `tidyverse` package and load the data. The data for this part of the practical is located in the data directory. Save the data in a variable called `project_dat`. Print the data to the screen to ensure it loaded correctly. -->

```{r}
#install.packages(tidyverse)
library(tidyverse)
setwd("../../")
project_dat<-read_csv("data/FossilAnts.csv")




```

2. (5pts) Check the datatypes of each column. There is a column called reference number. This is a static identifier - it should not be changed, and is an indexer used to identify this specimen uniquely. Do we want to treat it as an integer (this is an opinion question - answers may vary).

```
# We should not treat it as an integer as it represents a specific record. It is analogous with the record's "name." So, it should be a string.
```

3. (5 pts) Change the datatype of the reference number column to character. Take a peek at the function as.character().

```{r}
project_dat$reference_no<-as.character(project_dat$reference_no)


```

4. (5 pts) Look at your data. What are the missing data values? In particular, have a look at the `Tribe` column. In your opinion, are these intelligent missing values for the dataset? Why or why not? If not, how would you like to change them? 

```
#The missing data values in the Tribe column are "none" and "NA." It seems as though none means the researcher did not find the tribe for the genus, and NA means they did not input a value or that a tribe was not applicable. In one row, a Camponotini was given and a genus was not. The the very next row, camponotites was given for genus and the tribe was none. This and the fact that most of the column is "none" leads me to believe that the column tribe is irrelevant. I would move the terms to the genus column and then remove the tribe column.

```

6. (5 points) In the surveys dataset, we have genus and species split between two columns. Here, we have them combined. What are the pros and cons of the way we have recorded taxa in these two data sets.

```
# The pros are that it is easier to find the exact species name by using the search function. A con is that you cannot just pull up a genus from the species column. However, since a genus column was also given, you can do this by searching the genus column. 
```

7. (5pts) Please look at the help page for the `separate` function.

```{r}
?separate
```


7. (10 pts) How could you separate one column into two? 

```{r}
column_separated<-separate(data=project_dat, col=taxon, into=c("species", "genus"), sep="_", fill="warn")
```

Next, we will test a hypothesis. Your hypothesis is that there are more specimens in the 75 million years ago (mya) - 100 mya interval than the 30 mya to the present interval.

8. (5 pts) Write out the steps you would take to address this question. Will you need to split up the data? Will you need to group the data based on the values in some column? 
You will need to filter out the time intervals. However, I did not need to split the columns. I assigned the filter for the first interval and counted. Then, I did the same for the second. 


9. (15 pts) Perform the operations you described in (8). 

```{r}
first_interval<-project_dat %>% 
filter(min_ma>=75, max_ma<=100)

first_interval %>% 
count()
#48 in the first interval (75-100 mya)

second_interval<-project_dat %>% 
filter(min_ma>=0, max_ma<=30)

second_interval %>% 
count()
#313 in the second interval (present to 30 mya)

```

10. (5 pts) Do the results of your code support the hypothesis? 

```
#No, the results do not support the hypothesis. 
```


11. (10 pts) Save the dataframe with the split taxon columns into a new directory called `project_one_data_output`. Save it as a csv file called "column_separated.csv" 

```{r}

write_csv(column_separated,"column_separated.csv")


```

12. (10 pts) Produce a histogram of the number of specimens by minimum age. Look at the geom_histogram() help.

```{r}
ggplot(project_dat, aes(min_ma))+geom_histogram()


```

13. (5 pts) Change the font size on your histogram so that it can be read comfortably on your computer screen from five feet away. Save the file as "large_font.pdf".

```{r}
large_font<-ggplot(project_dat, aes(min_ma))+geom_histogram()
large_font<-large_font+theme(text=element_text(size=30))
ggsave("large_font.pdf", large_font)

```

14. Finally, produce a histogram of counts for each subfamily.

```{r}
ggplot(column_separated, aes(subfamily)) + geom_histogram(stat="count")


```


# Graduate Students

1. You will do the above steps with the classroom dataset.

2. Next, you will choose three data steps from above to perform with your own data. This could be grouping, mutating, cleaning NA values, visualization, or any combination of the three. 
```{r}
GM<-read.csv("../GM.csv")
#AMW: Fixed path for you
GM<-GM %>% 
  na.omit()

SL_box<-ggplot(data=GM, mapping=aes(x=Sex, y=SL))+geom_boxplot()+labs(x="Sex",y="Standard Length in mm")
LL_box<-ggplot(data=GM, mapping=aes(x=Sex, y=LL))+geom_boxplot()+labs(x="Sex", y="Lower Lateral Line Scale Count")
TV_box<-ggplot(data=GM, mapping=aes(x=Sex, y=TV))+geom_boxplot()+labs(x="Sex", y="Total Length in mm")
CPD_box<-ggplot(data=GM, mapping=aes(x=Sex, y=CPD))+geom_boxplot()+labs(x="Sex", y="Caudal Peduncle Scale Count")

graph1<-grid.arrange(SL_box, LL_box, TV_box, CPD_box, top="E. gutselli meristics by sex")
```
3. Why did you choose these steps? What function does this serve for your thesis and/or other research projects? 

First, I chose to remove NAs. This is an important function for my research projects as I may need to calculate values in the future. I chose to visualize my data. I made 4 plots showing the different meristics by sex. Then, I arranged the graphs of meristic values by sex in order to compare the differences of the species by sex. 

#AMW: Are these the results you expected - that some are basically the same, and some aren't?

